---
title: "Learning models"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = "figs/",
  fig.width = 7.252,
  fig.height = 4,
  comment = "#>",
  fig.retina = 3
)

# Load libraries, functions, and dir paths
source(here::here('code', '0setup.R'))

# Load formatted data
data <- readRDS(dir$data_formatted)

# Select data
df_us <- data$hist_us
df_china <- data$hist_china
df_germany <- data$hist_germany
```

## Two-factor learning model

A two-factor learning model relating the unit cost, $c_{it}$, in year $t$ and country $i$ of solar PV modules to the cumulative installed PV capacity in year $t$, $q_t$, and globally-averaged polysilicon prices in year $t$, $s_t$, (the primary input material to PV modules) is given by:

$$
c_{it} = 
\alpha_i q_t^{\beta_i} s_t^{\gamma_i}
$$

Here, $\alpha_i$ is a constant related to the starting year conditions in country $i$, $\gamma_i$ measures the sensitivity to polysilicon prices, and $\beta_i$ is the learning coefficient in country $i$. A log transformation produces the typical log-log formulation of the model that can be estimated via ordinary linear regression:

$$
\ln c_{it} = 
\ln \alpha_i + 
\beta_i \ln q_t + 
\gamma_i \ln s_t 
$$

For each country, $i$, we estimate "global learning" coefficients (Extended Data Table 1), $i$, using linear least-squares regression on (2). The "global learning" model is based on the assumptions that 1) learning was global during the historical period of scaling up PV module production, and 2) variations in country-level module pricing were due to transportation, administrative, and other non-learning costs. We then construct counterfactual "national learning" scenarios by assuming that the learning-related cost improvements in country $i$ from the starting year, $t_0$, are only derived from incrementally more nationally-installed PV capacity: 

$$
q_t - q_{t-1} = q_{t_0} + (q_{it} - q_{it-1}) + (1 - \lambda_i) (q_{jt} - q_{jt-1}),
$$

\noindent
where $q_{it}$ is the cumulative installed capacity in country $i$ in year $t$, $q_{jt}$ is the cumulative installed capacity in all other countries in year $t$, and $\lambda_i$ is a value ranging from 0 to 1. This defines a scenario of incremental capacity in each year that increasingly comes from national as opposed to global installations as $\lambda_i$ shifts from 0 to 1. Since $0 \le \lambda_i \le 1$, equation (2) can be iteratively estimated using every possible value of $\lambda_i$ from $0$ to $1$. The "best" value of $\lambda_i$ for country $i$ is the one that results in the smallest sum of squared residuals.

## Interpreting $\lambda$ in the context of learning

A core assumption in the above model is that national and global learning are proportionally derived from national and global cumulative installed capacities. With this in mind, the value of $\lambda_i$ itself cannot be directly interpreted as the proportion of national learning; rather, the proportion of national learning must be computed as the ratio between the national and global installed capacities for a given value of $\lambda_i$:

$$
\begin{aligned}
p_{it} &= \frac{q_{it}}{q_t} \\
    &= \frac{q_{it}}{q_{it} + (1 - \lambda_i) q_{jt}}
\end{aligned}
$$

Using this definition, 100% of learning will be national when $\lambda_i = 1$; however, if $\lambda_i = 0$, then national learning will be $q_{it} / q_t$, which will always be greater than 0. The figure below illustrates the relationship between $\lambda_i$ and the median value of $p_i$ in each year for each country.

```{r, echo=FALSE}
compareCapData <- function(df) {
    lambda <- seq(0, 1, 0.2)
    q0 <- df$cumCapKw_world[1]
    results <- list()
    for (i in 1:length(lambda)) {
        results[[i]] <- df %>% 
            mutate(
                q = cumsum(q0 + annCapKw_nation + (1 - lambda[i]) * annCapKw_other),
                q_i = cumsum(q0 + annCapKw_nation), 
                p = q_i / q, 
                median_p = round(median(p), 2)
            )
    }
    result <- do.call(rbind, results)
    result$lambda <- rep(lambda, each = nrow(df))
    result <- result %>% 
        mutate(label = paste0("λ = ", lambda, ", p_i = ", median_p))
    return(result)
}
```

```{r, echo=FALSE, fig.width=12, fig.height=3.5}
plotData <- rbind(
    compareCapData(df_us) %>% mutate(country = "U.S."),
    compareCapData(df_china) %>% mutate(country = "China"),
    compareCapData(df_germany) %>% mutate(country = "Germany")
) 

ggplot(plotData) + 
    geom_line(
        aes(x = year, y = q, group = lambda)
    ) + 
    geom_text(
        data = plotData %>% 
            group_by(country, lambda) %>% 
            filter(year == max(year)),
        aes(x = year, y = q, label = label),
        hjust = 0, nudge_x = 0.2
    ) +
    facet_wrap(vars(country)) +
    expand_limits(x = c(2006, 2030)) +
    theme_bw() + 
    labs(
        x = "Year", 
        y = "Cumulative installed capacity (kW)"
    )

labelData <- plotData %>% 
    filter(lambda == 0.4) %>% 
    distinct(country, median_p)
p_us <- labelData %>% filter(country == "U.S.") %>% pull(median_p)
p_china <- labelData %>% filter(country == "China") %>% pull(median_p)
p_germany <- labelData %>% filter(country == "Germany") %>% pull(median_p)
```

As the figure shows, the same value of $\lambda_i$ does not translate to the same proportion of national learning for each country. For example, if $\lambda_i = 0.4$, then the proportion of national learning for each country is:

- U.S.: `r percent(p_us)`
- China: `r percent(p_china)`
- Germany: `r percent(p_germany)`

## When should we use $\lambda_i$? 

The hybrid model that includes a $\lambda_i$ term is rooted in the concept that capacity installed within a nation should contribute to that nation's learning rate. However, if a nation largely _imports_ the PV panels that are installed domestically, then the learning is more likely occurring in the country where those panels were _manufactured_ rather than in the country where they were installed.

This raises a few questions.

### Perhaps we should _not_ use the hybrid model for the "global markets" scenario?

One benefit of estimating the hybrid model for the "global markets" scenario is that it produces a value for $\lambda_i$ that can be used as a starting value in the national markets scenario. However, it also attributes national learning to domestic capacity that was largely imported. 

If we instead stick with a traditional learning curve using global capacities for the "global markets" scenario, then we're no longer attributing learning as global versus national. We're simply saying this was the historical learning rate in country $i$. It's only in the "national markets" scenario where we have to then start to break down how much learning is coming from national versus global capacity. That said, in this counter factual world the domestic capacity is being domestically manufactured, so the hybrid model makes sense. The drawback here is that we now have to pick values both starting and end values for $\lambda_i$ as well as how we want them to change over time.

### Perhaps we should consider adding production into the model? 

I'm wondering if a three-factor model might make sense where we include both installed capacity as well as domestic production somehow?


